    \chapter{Feasibility Study}
        \section{Technical Feasibility}

The technical feasibility of a lip-reading application falls on usage of advanced image and video processing techniques in order to capture and process and analyze clear lip movement. The system should be seamlessly running with speech-to-text capability. And usage of natural language processing is a must in order to increase the transcription ability.  Real-time processing capability and consideration of hardware requirements is a must for bringing the system into practical use. Through the use of large and varied dataset using an effective model and also keeping the functioning of system in different lighting conditions in mind a consistent user experience is expected.

\section{Economic Feasibility}

The economic feasibility of a project verifies project's financial viability by examining a project's costs, benefits, and risks to determine whether it is financially viable and worthwhile to pursue.
For this system, the economic feasibility would involve the cost of training and fine-tuning multiple image and video processing models, implementation of the project its software development keeping required hardware in mind along with its maintenance cost.

\section{Schedule Feasibility}

The scheduling feasibility of a project is an assessment of whether the project and be completed within a specified time frame maintaining quality standard. There are several factors which could impact the schedule, including the availability of the resource materials the project is estimated to take a little over then 3 months. There is expectations to finish the documentation and testing of the system in the specified time frame.