\pagenumbering{roman}
% \newgeometry{right = 1in}
% \begingroup
% \RaggedRight
%         \large
%         \chapter*{Certificate of Approval}
% 		\normalsize 
    
%     This is to certify that this minor project work entitled \textbf{LIP READING USING CONVOLUTIONAL NEURAL NETWORKS} submitted by Chirag Khatiwada (KCE077BCT001), Bishesh Pokharel (KCE077BCT014), Mahim Rawal (KCE077BCT019) and Rowel Maharjan (KCE077BCT027),  has been examined and accepted for the partial fulfillment of the requirements for the degree of Bachelor's in Computer Engineering.

%     \vspace{1in}
%     \begin{multicols}{2}
        
%         \begin{center}
%             ..........................................\\
%             \textbf{Er. Dr. Kishor Kumar Adhikari}\\
%             External Examiner \\
%             Assistant Professor,\\ 
%             Department of Electronics and Computer Engineering,\\
%             National College of Engineering
%         \end{center}
%     \columnbreak
%         \begin{center}
%          ..........................................................\\
%         \textbf{Er.Mukesh Kumar Pokharel}\\
%     Project Supervisor \\
%     Assistant Lecturer,\\
%     Department of Computer Engineering,
%     Khwopa College of Engineering
%         \end{center}
        
        
        
%     \end{multicols}
    
   % \vspace{0.7in}

%     \begin{center}
        
%           .........................................\\
%    \textbf{Er. Dinesh Gothe}\\
%    Head of Department\\
%    Department of Computer Engineering,\\
%    Khwopa College of Engineering\\
 
%        \end{center}
%    \pagebreak
% \endgroup
% Specify the length of whitespace section in inches
\newlength{\whitespaceLength}
\setlength{\whitespaceLength}{1in} % Adjust the value as needed

% Use \vspace to leave the whitespace section
\vspace*{\whitespaceLength}

% Your content here after the whitespace section
\begin{center}
\large{\textbf{CERTIFICATE OF APPROVAL}}
\end{center}

\begingroup
 		\addcontentsline{toc}{section}{Certificate of Approval}
\RaggedRight
            \vspace{0.5cm}
            This is to certify that this minor project work entitled \textbf{"LIP READING USING CONVOLUTIONAL NEURAL NETWORKS"} submitted by Chirag Khatiwada (KCE077BCT001), Bishesh Pokharel (KCE076BCT014), Mahim Rawal (KCE077BCT019) and Rowel Maharjan (KCE077BCT027),  has been examined and accepted as the partial fulfillment of the requirements for the degree of Bachelor in Computer.\\
            \vspace{0.5in}
            \begin{multicols}{2}
            	\begin{center}
            	
                \setlength{\whitespaceLength}{0.5in}
                \null\vspace{\whitespaceLength}
            		 
          .........................................................\\
            \textbf{Er. Kishor Kumar Adhikari, PhD}\\
            External Examiner,\\
            Associate Professor,\\ 
            Department of Electronics and Computer Engineering,\\
            National College of Engineering
            		
            	\end{center}
            
            \columnbreak
            	\begin{center}
                \setlength{\whitespaceLength}{0.5in}
                \null\vspace{\whitespaceLength}
         ..........................................................\\
        \textbf{Er.Mukesh Kumar Pokharel}\\
    Project Supervisor \\
    Assistant Lecturer,\\
    Department of Computer Engineering,
    Khwopa College of Engineering
        \end{center}
            \end{multicols}
            \vspace{0.5in}
            \begin{center}
            	..........................................\\
            	\textbf{Er. Dinesh Gothe}\\
            	Head of Department,\\
            	Department of Computer Engineering\\
            	Khwopa College of Engineering
            \end{center}
            \pagebreak
            \endgroup
		\addcontentsline{toc}{section}{Copyright}
		\large
			\chapter*{Copyright}
		\normalsize
		
		The authors have agreed that the library, Department of Computer Engineering, Khwopa College of Engineering may make this project report freely available
for inspection. Moreover, the author has agreed that permission for extensive copying of
this project report for scholarly purposes may be granted by the professor(s) who
supervised the work recorded herein or, in their absence, by the Head of the Department
wherein the thesis was done. It is understood that recognition will be given to the
authors of this project report and the Department of Computer Engineering, Khwopa College of Engineering, for any use of the material of this project
report. Copying, publication, or the other use of this project report for financial gain
without the approval of the Department of Computer Engineering, Khwopa College of Engineering, and the author's written permission is prohibited.\\ \\
Request for permission to copy or to make any other use of this project report in whole or in part should be addressed to:
\newline
\newline
\newline
\newline
Head of Department\\
Department of Computer Engineering\\
Khwopa College of Engineering\\
Libali, Bhaktapur \\
Nepal
		\break

	\newpage
	    \addcontentsline{toc}{section}{Acknowledgement}
		 \large
            \chapter*{Acknowledgement}
            % \addcontentsline{toc}{chapter}{Acknowledgement}
            \normalsize
            
            We would like to express our heartfelt gratitude to our HOD, \textbf{Er. Dinesh Gothe}, for providing us with the opportunity and the confidence to act on our will to work on this project and proposal. Also, not forgetting the great help of our supervisor, \textbf{Er. Mukesh Kumar Pokharel}, for providing us with the vision to work on our project.


            \vspace{1 cm}


            \begin{tabular}{ll}
                \centering
                \textbf{Chirag Khatiwada} & KCE077BCT001 \\
                \textbf{Bishesh Pokharel} & KCE077BCT014 \\
                \textbf{Mahim Rawal} & KCE077BCT019 \\
                \textbf{Rowel Maharjan} & KCE077BCT027 \\
            \end{tabular}

                   

   
		\break
		
		\addcontentsline{toc}{section}{Abstract}
		\large
			\chapter*{Abstract}
		\normalsize
			Decoding text from a speaker's facial movements, or lip reading, is a skill that has great potential for people with speech or hearing impairments. In addition to improving communication, it makes captioning easier in difficult audio situations, such as crowded spaces or far-off speakers. Automated lip reading systems have been made possible by recent advances in computer vision and machine learning. In order to improve the accuracy of visual voice recognition, this research project suggests a hybrid architecture that combines Bidirectional Long Short-Term Memory(BiLSTM) and Convolutional Neural Networks(CNN), taking advantage of their respective advantages. To ensure robustness and generalization, the hybrid model will be trained on an extensive dataset that spans a variety of speakers, languages, and environmental situations. Model parameters will be optimized using fine-tuning processes to provide adaptability in a variety of lip reading circumstances. In order to simplify the implementation process, the project is limited only to visual data. 
            \break	\break
            \textbf{Keywords}: Lip reading, Convolutional Neural Networks (CNN),
            Bidirectional Long Short-Term Memory (BiLSTM) networks

		\break


		\tableofcontents
		
		\addcontentsline{toc}{section}{List of Tables}
		\listoftables
		\break
		\pagebreak

		\addcontentsline{toc}{section}{List of Figures}
		\listoffigures
		\break
	
	
	
		\addcontentsline{toc}{section}{List of Symbols and Abbreviation}
		\Large
			\begingroup
				\let\clearpage\relax
				\chapter*{List of Symbols and Abbreviation}
			\endgroup
		\normalsize

            \begin{tabular}{p{2cm} p{10cm}}
                ADAM & Adaptive Moment Estimation\\
                AL & Automatic Lipreading \\
                ASR & Automatic Speech Recognition \\
                Bi-LSTM & Bi-directional Long Short Term Memory\\
                CNN & Convolutional Neural Networks \\
                GB & Giga Byte\\
                GCP & Google Cloud Platform\\
                GPU & Graphics Processing Unit\\
                LSTM & Long Short Term Memory\\
                RNN & Recurrent Neural Network\\
                VM & Virtual Machine\\
                SAT & Speaker Adaptive Training \\

            \end{tabular}
		\break
		\pagebreak